---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```


# sjSDM - Fast and accurate Joint Species Distribution Modeling

[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
![R-CMD-check](https://github.com/TheoreticalEcology/s-jSDM/workflows/R-CMD-check/badge.svg?branch=master)
[![Publication](https://img.shields.io/badge/Publication-10.1111/2041-green.svg)](https://besjournals.onlinelibrary.wiley.com/doi/abs/10.1111/2041-210X.13687)

## Overview

A scalable method to estimates joint Species Distribution Models (jSDMs) based on the multivariate probit model through Monte-Carlo approximation of the joint likelihood. The numerical approximation is based on 'PyTorch' and 'reticulate', and can be calculated on CPUs and GPUs alike. 

The method is described in Pichler & Hartig (2021) <doi.org/10.1111/2041-210X.13687> A new joint species distribution model for faster and more accurate inference of species associations from big community data.

The package includes options to fit various different (j)SDM models:

* jSDMs with Binomial, Poisson and Normal distributed responses
* jSDMs based on deep neural networks
* Spatial auto-correlation can be accounted for by spatial eigenvectors or trend surface polynomials 

To get more information, install the package and run

```{r, eval=FALSE}
library(sjSDM)
?sjSDM
vignette("sjSDM", package="sjSDM")
```

## Installation

__sjSDM__ is based on 'PyTorch', a 'python' library, and thus requires 'python' dependencies. The 'python' dependencies can be automatically installed by running:

```{r,eval=FALSE}
library(sjSDM)
install_sjSDM()
```

If this didn't work, please check the troubleshooting guide:
```{r,eval=FALSE}
library(sjSDM)
?installation_help
```

## Usage

Let's first simulate a community data set:

```{r}
library(sjSDM)
set.seed(42)
community <- simulate_SDM(sites = 100, species = 10, env = 3, se = TRUE)
Env <- community$env_weights
Occ <- community$response
SP <- matrix(rnorm(200, 0, 0.3), 100, 2) # spatial coordinates (no effect on species occurences)
```

Estimate jSDM:

```{r}
model <- sjSDM(Y = Occ, env = linear(data = Env, formula = ~X1), spatial = linear(data = SP, formula = ~0+X1:X2), se = TRUE, family=binomial("probit"), sampling = 100L)
summary(model)
```

Update model (change main effects to quadratic effects):

```{r}
model2 <- update(model, env_formula = ~I(X1^2))
summary(model2)
```

Change linear part of model to a deep neural network:

```{r}
DNN <- sjSDM(Y = Occ, env = DNN(data = Env, formula = ~.), spatial = linear(data = SP, formula = ~0+X1:X2), se = TRUE, family=binomial("probit"), sampling = 100L)
summary(DNN)
```
