---
title: "sjSDM: Help with the installation of dependencies"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{sjSDM: Help with the installation of dependencies}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "This document explains how to install dependencies for the sjSDM package. \n \n \n"
editor_options: 
  
  chunk_output_type: console
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = FALSE, eval = FALSE)
```

The `r knitr::inline_expr('sjSDM::install_sjSDM()')` function can install automatically all necessary 'python' dependencies but it can fail sometimes because of individual system settings or if other 'python'/'conda' installations get into the way. 

## PyTorch Installation - Before you start:

A few notes before you start with the installation (skip this point if you 
do not know conda):

* existing 'conda' installations: make sure you have the latest 'conda' or 'miniconda' version and remove unnecessary 'conda' installations.
* existing 'conda' or 'virtualenv' environments (skip this point if you do not know 'conda'): we currently enforce the usage of a specific environment called 'r-sjsdm',so if you want use a custom environment it should be named 'r-sjsdm'
 
 
## Windows - automatic installation:
 
Sometimes the automatic 'miniconda' installation (via 
`r knitr::inline_expr('sjSDM::install_sjSDM()')` ).doesn't work because of white spaces in the user's name. But you can easily download and install 'conda' on your own:

Download and install the latest ['conda' version](https://www.anaconda.com/products/individual)

Afterwards run:
```{r,eval=FALSE}
install_sjSDM(version = c("gpu")) # or "cpu" if you do not have a proper gpu device
```

 
Reload the package and run the example, if this doesn't work:

* Restart RStudio
* Install manually pytorch, see the following section
 
  
## Windows - manual installation:
 
Download and install the latest ['conda' version](https://www.anaconda.com/products/individual)

Open the command window (cmd.exe - hit windows key + r and write cmd)

Run in cmd.exe:

```{bash, eval=FALSE}
conda create --name r-sjsdm python=3.7
conda activate r-sjsdm
conda install pytorch torchvision cpuonly -c pytorch # cpu
conda install pytorch torchvision cudatoolkit=11.3 -c pytorch #gpu
python -m pip install pyro-ppl torch_optimizer madgrad
```
 
 
Restart R, try to run the example, and if this doesn't work:

* Restart RStudio
* See the 'Help and bugs' section
 
    
## Linux - automatic installation:
 
Run in R: 
```{r,eval=FALSE}
install_sjSDM(version = c("gpu")) # or "cpu" if you do not have a proper gpu device
```

Restart R try to run the example, if this doesn't work:

* Restart RStudio
* Install manually 'PyTorch', see the following section

  
  
## Linux - manual installation:
 
We strongly advise to use a 'conda' environment but a virtual environment should also work. The only requirement is that it is named 'r-sjsdm'

Download and install the latest ['conda' version](https://www.anaconda.com/products/individual)

Open your terminal and run:
```{bash, eval=FALSE}
conda create --name r-sjsdm python=3.7
conda activate r-sjsdm
conda install pytorch torchvision cpuonly -c pytorch # cpu
conda install pytorch torchvision cudatoolkit=11.3 -c pytorch #gpu
python -m pip install pyro-ppl torch_optimizer madgrad
```
 
 
Restart R try to run the example, if this doesn't work:

* Restart RStudio
* See the 'Help and bugs' section
 
 
## MacOS - automatic installation:
 
Run in R: 
```{r,eval=FALSE}
install_sjSDM()
```

Restart R try to run the example, if this doesn't work:

* Restart RStudio
* Install manually 'PyTorch', see the following section

 
 
## MacOS - manual installation:
 
We strongly advise to use a 'conda' environment but a virtual environment should also work. The only requirement is that it is named 'r-sjsdm'

Download and install the latest conda [conda version](https://www.anaconda.com/products/individual)

Open your terminal and run:
```{bash, eval=FALSE}
conda create --name r-sjsdm python=3.7
conda activate r-sjsdm
python -m pip install torch torchvision torchaudio 
python -m pip install pyro-ppl torch_optimizer madgrad
```
 
 
Restart R try to run the example, if this doesn't work:

* Restart RStudio
* See the 'Help and bugs' section

## Help and bugs

* ask for help via the [issue tracker](https://github.com/TheoreticalEcology/s-jSDM/issues). Please post the output of  `r knitr::inline_expr('sjSDM::install_diagnostic()')`as a quote. 
* report bugs with a [reproducible example](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-emple)
 via the [issue tracker](https://github.com/TheoreticalEcology/s-jSDM/issues)



## Example
```{r,eval=FALSE}
library(sjSDM)
community <- simulate_SDM(sites = 100, species = 10, env = 5)
Env <- community$env_weights
Occ <- community$response

model <- sjSDM(X = Env, Y = Occ, formula = ~0+X1*X2 + X3 + X4)
summary(model)
```