---
title: "sjSDM: help on installing dependencies"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Dependencies}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "This document explains how to install dependencies for the sjSDM package. \n \n \n"
editor_options: 
  chunk_output_type: console
  
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = F)
```




# Before you start
A few notes before you start with the installation:

- existing conda installations (skip this point if you do not know conda): make sure you have the latest conda3/miniconda3 installation and remove unneccessary conda installations.
- existing conda/virtualenv (skip this point if you do not know conda): the R package reticulate has problems to find the right environment and it is better to remove unneccessary environments (see the [troubleshooting](#troubleshooting) section




# Windows - automatic installation 
Conda is the easiest way to install python and python packages on windows:
Install the latest [conda version](https://www.anaconda.com/distribution/)

Afterwards, run in R:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
install_sjSDM(version = c("gpu")) # or "cpu" if you do not have a proper gpu device
```
Try to load the package and run the example from ?sjSDM, if this doesn't work:

- Restart RStudio
- Install manually pytorch, see the following section


 
# Windows - manual installation {#ww}
Conda is the easiest way to install python and python packages on windows:

* Install the latest [conda version](https://www.anaconda.com/distribution/)
* Open the command window (cmd.exe - hit windows key + r and write cmd)

Run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
$ conda install pytorch torchvision cudatoolkit=10.1 -c pytorch #gpu
```

Start R and run:
Download, install, and test the package:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
community <- simulate_SDM(sites = 100, species = 10, env = 5)
Env <- community$env_weights
Occ <- community$response

model <- sjSDM(X = Env, Y = Occ, formula = ~0+X1*X2 + X3 + X4)
summary(model)
```
Try to load the package and run the [example](#example), if this doesn't work:

* Restart RStudio
* See the [troubleshooting](#troubleshooting) section




# Linux - automatic installation
Run in R:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
install_sjSDM(version = c("gpu")) # or "cpu" if you do not have a proper gpu device
```
Try to load the package and run the [example](#example), if this doesn't work:

* Restart RStudio
* Install manually pytorch, see the following section





# Linux - manual installation {#lm}
## Via pip
python3 is pre-installed on most linux distributions, but you have to check that the minimal requirement of python >= 3.6 is met: 

```
$ python3 --version 
$ python --version
```

Check if pip is installed:
```
$ pip3 --version 
$ pip --version
```
If not then install pip:
```
$ sudo apt install python3-pip # for ubuntu/deb   # e.g. for ubuntu
```

It depends on your ENVIRONMENTAL variables whether pip or pip3 corresponds to python3. You can check the python version of pip via:
```
$ pip --version
```

In my case, I have to use "pip3":

```
$ pip3 install torch==1.4.0+cpu torchvision==0.5.0+cpu -f https://download.pytorch.org/whl/torch_stable.html #cpu
$ pip3 install torch torchvision #gpu
```

Run in R:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
```
Try to load the package and run the [example](#example), if this doesn't work:

* Restart RStudio
* See the [troubleshooting](#troubleshooting) section


## Via virtualenv
Create a virtualenv and install dependencies:
```
$ python3 -m pip install --user virtualenv
$ python3 -m venv ~/sjSDM_env
$ source ~/sjSDM_env/bin/activate
$ pip install torch==1.4.0+cpu torchvision==0.5.0+cpu -f https://download.pytorch.org/whl/torch_stable.html #cpu
$ pip install torch torchvision #gpu
```

start R from within the environment and run:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
```
Try to load the package and run the [example](#example), if this doesn't work:

* Restart R/RStudio
* See the [troubleshooting](#troubleshooting) section


## Via conda
Install the latest [conda version](https://www.anaconda.com/distribution/) and run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
$ conda install pytorch torchvision cudatoolkit=10.1 -c pytorch #gpu
```
start RStudio from  within the conda env and run:
Try to load the package and run the [example](#example), if this doesn't work:

* Restart R/RStudio
* See the [troubleshooting](#troubleshooting) section




# MacOS
## Conda
Install the latest [conda version](https://www.anaconda.com/distribution/) and run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
```
start R and run:
```{r,eval=FALSE}
devtools::install_github("https://github.com/TheoreticalEcology/s-jSDM", subdir = "sjSDM")
library(sjSDM)
```
Try to load the package and run the [example](#example), if this doesn't work:

* Restart R/RStudio
* See the [troubleshooting](#troubleshooting) section

For GPU support on MacOS, you have to install the cuda binaries yourself, see [PyTorch for help](https://pytorch.org/)


# Troubleshooting
After several installation attempts you might have corrupt and unneccessary conda environments which we have to remove:
Open the shell (Windows users: run cmd.exe):
```
$ conda env list
```
will list the paths to all your environments, except for the "base" env, delete them via:
```
$ rm - r <path-to-env1>
$ rm - r <path-to-env2>
```
Then continue with the manual install instructions:

1. [Windows](#ww)
2. [Linux](#lw)
3. [MacOS](#macos)





# Example
```{r,eval=FALSE}
library(sjSDM)
community <- simulate_SDM(sites = 100, species = 10, env = 5)
Env <- community$env_weights
Occ <- community$response

model <- sjSDM(X = Env, Y = Occ, formula = ~0+X1*X2 + X3 + X4)
summary(model)
```