---
title: "sjSDM: help on installing dependencies"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Dependencies}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "This document explains how to install dependencies for the sjSDM package. \n \n \n"
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = F)
```





# Windows

Conda is the easiest way to install python and python packages on windows:
- Install the latest [conda version](https://www.anaconda.com/distribution/)
- Open the command window (cmd.exe - hit windows key + r and write cmd)

Run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pip
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
$ conda install pytorch torchvision cudatoolkit=10.1 -c pytorch #gpu
```



# Linux

## Pip
python3 is pre-installed on most linux distributions, but you have to check that the minimal requirement of python >= 3.6 is met: 

```
$ python3 --version 
$ python --version
```

Install pip
```
$ sudo apt install python3-pip # for ubuntu/deb
```

Create a virtualenv and install dependencies:
```
$ python3 -m pip install --user virtualenv
$ python3 -m venv ~/sjSDM_env
$ source ~/sjSDM_env/bin/activate
$ pip install torch==1.4.0+cpu torchvision==0.5.0+cpu -f https://download.pytorch.org/whl/torch_stable.html #cpu
$ pip install torch torchvision #gpu
```

start RStudio from within the virtualenv and try the [example](#example)

## Conda

Install the latest [conda version](https://www.anaconda.com/distribution/) and run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pip
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
$ conda install pytorch torchvision cudatoolkit=10.1 -c pytorch #gpu
```
start RStudio from  within the conda env and try the [example](#example)

# MacOS

## Conda
Install the latest [conda version](https://www.anaconda.com/distribution/) and run:
```
$ conda create --name sjSDM_env python=3.7
$ conda activate sjSDM_env
$ conda install pip
$ conda install pytorch torchvision cpuonly -c pytorch # cpu
```
start RStudio from  within the conda env and try the [example](#example)

For GPU support on MacOS, you have to install the cuda binaries yourself, see [PyTorch for help](https://pytorch.org/)
