---
title: "Getting started with sjSDM: a scalable joint Species Distribution Model"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 

  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with sjSDM: a scalable joint Species Distribution Model}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "sjSDM is a scalable joint species distribution model based on Monte-Carlo intergration of the joint likelihood. \n \n \n"
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = FALSE)
```


# Getting started 

Load the package via 

```{r,eval=FALSE}
library(sjSDM)
```


## Installing dependencies

sjSDM depends on Anaconda which will be installed automatically if it is not available.

To install the CPU version, run:
```{r, eval = F, purl=FALSE}
install_sjSDM()
```

For the GPU version (only NVIDIA GPUs are supported), run:
```{r, eval = F, purl=FALSE}
install_sjSDM(method = "gpu")
```


More details on and trouble-shooting for installing the necessary dependencies is explained in the vignette dependencies, call 

```{r,eval = F, purl=FALSE}
vignette("Dependencies", package = "sjSDM")
```

## Citing sjSDM

To cite sjSDM in a publication, use 

```{r,eval=FALSE}
citation("sjSDM")
```


# Working with sjSDM

sjSDM has a function to create test data: We create a dataset with 3 environmental predictors, 5 species and 100 sites.

```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
```


# Fitting a model

The model is then fit with the function sjSDM(). You have to provide predictors (can be also a data.frame) and response as matrices. 

```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = com$env_weights, iter = 100L, se=TRUE)
```


## Interpreting model output

Things you can do with the model output

```{r,eval=FALSE}
coef(model)
summary(model)
```

### Get species-species association matrix

```{r}
association = getCor(model)
```


## Adding quadratic predictors and interactions

sjSDM supports the R formula syntax.

E.g. interaction with intercept (if you want/need p-values, you have to set se = TRUE)
```{r,eval=FALSE}
model = sjSDM(Y = com$response, 
              env = linear(data = com$env_weights, formula = ~ X1*X2,), iter = 50L, se = TRUE)
summary(model)

```

E.g. quadratic effect without intercept:
```{r,eval=FALSE}
model = update(model, env_formula = ~0+ I(X1^2)+X1)
summary(model)
```


# Handling spatial autocorrelation
## Simulating spatial data
jSDMs account for correlation between species within communities (sites), in real datasets, however, communities (sites) are often also correlated (== spatial autocorrelation). Usually conditional autoregressive (CAR) models are used to account for the spatial autocorrelation in the residuals, which we, however, do not support yet. A similar approach is to condition the model on space, which we can do by using space as predictors. 

Let's first simulate test data:

1) Simulate jSDM without a link (normal response)

```{r,eval=FALSE}

com = simulate_SDM(env = 3L, species = 5L, sites = 100L, 
                   link = "identical", response = "identical")
X = com$env_weights
Y = com$response
```

2) add spatial residuals (create coordinates and use spatial distance matrix to draw autocorrelated residuals for each species)

```{r,eval=FALSE}
XYcoords = matrix(rnorm(200), 100, 2)+2
WW = as.matrix(dist(XYcoords))
spatialResiduals = mvtnorm::rmvnorm( 5L, sigma = exp(-WW))
```

3) Finish test data

```{r,eval=FALSE}
Ysp = Y + t(spatialResiduals)
Y = ifelse(Ysp < 0, 0, 1) # multivariate probit model

```

## Fitting spatial jSDM
There are three options to condition our model on space:

1) Trend surface model - linear

The idea of the trend surface model is to use the spatial coordinates within a polynom:

```{r,eval=FALSE}
colnames(XYcoords) = c("XX", "YY")
model = sjSDM(Y, env = linear(X, ~.), spatial = linear(XYcoords, ~0+XX+YY+XX:YY+I(XX^2)+I(YY^2)), iter = 100L)
summary(model)
```

2) Trend surface model - DNN

Sometimes a linear model and a polynom is not flexible enough to account for space. We can use a "simple" DNN for space to condition our linear environmental model on the space:

```{r,eval=FALSE}
colnames(XYcoords) = c("XX", "YY")
model = sjSDM(Y, env = linear(X, ~.), spatial = DNN(XYcoords, ~0+.), iter = 100L)
summary(model)
```


3) Using Moran's eigenvector map predictors


```{r,eval=FALSE}
SPeigen = generateSpatialEV(XYcoords)

model = sjSDM(Y, env = linear(X, ~.), spatial = linear(SPeigen, ~0+.), iter = 100L)
summary(model)
```


# Further analyses
## Anova
Type I analysis of variance for the two/three different components (environment, biotic associations, and space (optional)).
```{r,eval=FALSE}
an = anova(model)
plot(an)
```

## Internal meta-community structure
We can also visualize the internal meta-community structure (only for models with space):
```{r,eval=FALSE}
plot(an, internal = TRUE)
```



## Importance
We can also partition the effects of the different components. This function will be deprecated in future. 
Please use `plot(anova(model), internal = TRUE)`
```{r,eval=FALSE}
imp = importance(model)
plot(imp)
```



# Regularization
## Regularization on abiotic coefficients
sjSDM supports l1 (lasso) and l2 (ridge) regularization:
* alpha is the weighting between lasso and ridge
* alpha = 0.0 corresponds to pure lasso
* alpha = 1.0 corresponds to pure ridge
```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = linear(data = com$env_weights, formula = ~0+ I(X1^2),lambda = 0.5), iter = 50L)
summary(model)
```

## Regularization on species-species associations
We can do the same for the species associations:

```{r,eval=FALSE}
model = sjSDM(Y = com$response, 
              env = linear(data = com$env_weights, formula = ~0+ I(X1^2),lambda = 0.5),
              biotic = bioticStruct(lambda =0.1),
              iter = 50L)
summary(model)
```



## Regularization on the spatial model:

```{r,eval=FALSE}

model = sjSDM(Y, env = linear(X, ~X1+X2), spatial = linear(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```



# Fitting a non-linear (deep neural network) model 
```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
X = com$env_weights
Y = com$response

# three fully connected layers with relu as activation function
model = sjSDM(Y = Y, 
              env = DNN(data = X, formula = ~., hidden = c(10L, 10L, 10L), activation = "relu"), 
              iter = 50L, se = TRUE)
summary(model)
```



The methods for sjSDM() work also for the non-linear model:
```{r,eval=FALSE}
getCov(model) # species association matrix
pred = predict(model) # predict on fitted data
pred = predict(model, newdata = X) # predict on new data
```

Extract and set weights of model:
```{r,eval=FALSE}
weights = getWeights(model) # get layer weights and sigma
setWeights(model, weights)
```

Plot the training history:
```{r,eval=FALSE}
plot(model)
```


## We can also use a non-linear model as a spatial model:

```{r,eval=FALSE}

model = sjSDM(Y, env = linear(X, ~X1+X2), spatial = DNN(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```

## Or for both:
```{r,eval=FALSE}

model = sjSDM(Y, env = DNN(X, ~X1+X2), spatial = DNN(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```


# Fitting other distributions (e.g. species frequencies)
sjSDM supports other responses than presence-absence data:
Simulate non-presence-absence data:
```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L,link = "identical", response = "count") # samples from a Poisson distribution
X = com$env_weights
Y = com$response
```

## Poisson 
```{r,eval=FALSE}
model = sjSDM(Y, env = linear(X, ~.), se = TRUE, iter = 50L, family = poisson("log"))
summary(model)
```

## Normal (gaussian)
```{r,eval=FALSE}
model = sjSDM(log(Y+0.01), env = linear(X, ~.), se = TRUE, iter = 50L, family = gaussian("identity"))
summary(model)
```