---
title: "sjSDM: a scalable joint Species Distribution Model"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette for the sjSDM package}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "sjSDM is a scalable joint species distribution model based on Monte-Carlo intergration of the joint likelihood. \n \n \n"
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = F)
```


# Getting started 

Load the package via 

```{r,eval=FALSE}
library(sjSDM)
```


## Installing dependencies

sjSDM depends on the anaconda distribution of python and pytorch. You will get a warning if you don't have python or pytorch installed.

* anaconda needs to be installed by hand, outside R https://www.anaconda.com/distribution/
* once anaconda is installed, sjSDM can attempt to install pytorch automatically via 

```{r, eval = F}
install_sjSDM()
```

More details on and trouble-shooting for installing the neccessary dependencies is explained in the vignette dependencies, call 

```{r,eval = F}
vignette("Dependencies", package = "sjSDM")
```

## Citing sjSDM

To cite sjSDM in a publication, use 

```{r,eval=FALSE}
citation("sjSDM")
```


# Working with sjSDM

sjSDM has a function to create test data. Here, we create a dataset with 3 environmental predictors, 5 species and 100 sites.

```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
```

# Fitting a model

The model is then fit with the function sjSDM(). You have to provide predcitors (can be also a data.frame) and response as matrix. 

```{r,eval=FALSE}
model = sjSDM(X = com$env_weights, Y = com$response, iter = 10L)
```


## Interpreting model output

Things you can do with the model output

```{r,eval=FALSE}
coef(model)
summary(model)
getCov(model)
```

## Adding quadratic predictors and interactions

sjSDM supports formula description for the predictors.

E.g. interaction with intercept:
```{r,eval=FALSE}
model = sjSDM(X = com$env_weights, formula = ~ X1*X2,Y = com$response, iter = 10L)
summary(model)

```

E.g. quadratic effect without intercept:
```{r,eval=FALSE}
model = sjSDM(X = com$env_weights, formula = ~0+ I(X1^2),Y = com$response, iter = 10L)
summary(model)
```







