---
title: "sjSDM: a scalable joint Species Distribution Model"
author: "Maximilian Pichler & Florian Hartig, Theoretical Ecology, University of Regensburg"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Vignette for the sjSDM package}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
abstract: "sjSDM is a scalable joint species distribution model based on Monte-Carlo intergration of the joint likelihood. \n \n \n"
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F}
set.seed(123)
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4.5, fig.align='center', warning=FALSE, message=FALSE, cache = F)
```


# Getting started 

Load the package via 

```{r,eval=FALSE}
library(sjSDM)
```


## Installing dependencies

sjSDM depends on the anaconda distribution of python and pytorch. You will get a warning if you don't have python or pytorch installed.

* anaconda needs to be installed by hand, outside R https://www.anaconda.com/distribution/
* once anaconda is installed, sjSDM can attempt to install pytorch automatically via 

```{r, eval = F}
install_sjSDM()
```

More details on and trouble-shooting for installing the neccessary dependencies is explained in the vignette dependencies, call 

```{r,eval = F}
vignette("Dependencies", package = "sjSDM")
```

## Citing sjSDM

To cite sjSDM in a publication, use 

```{r,eval=FALSE}
citation("sjSDM")
```


# Working with sjSDM

sjSDM has a function to create test data. Here, we create a dataset with 3 environmental predictors, 5 species and 100 sites.

```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
```

# Fitting a model

The model is then fit with the function sjSDM(). You have to provide predcitors (can be also a data.frame) and response as matrix. 

```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = com$env_weights, iter = 10L)
```


## Interpreting model output

Things you can do with the model output

```{r,eval=FALSE}
coef(model)
summary(model)
getCov(model)
```

## Adding quadratic predictors and interactions

sjSDM supports formula description for the predictors.

E.g. interaction with intercept and calculate p-values:
```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = linear(data = com$env_weights, formula = ~ X1*X2,), iter = 50L, se = TRUE)
summary(model)

```

E.g. quadratic effect without intercept:
```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = linear(data = com$env_weights, formula = ~0+ I(X1^2)), iter = 50L, se = TRUE)
summary(model)
```


## Regularization
### Regularization on abiotic coefficients
sjSDM supports l1 (lasso) and l2 (ridge) regularization:
* alpha is the weighting between lasso and ridge
* alpha = 0.0 corresponds to pure lasso
* alpha = 1.0 corresponds to pure ridge
```{r,eval=FALSE}
model = sjSDM(Y = com$response, env = linear(data = com$env_weights, formula = ~0+ I(X1^2),lambda = 0.5), iter = 50L)
summary(model)
```

### Regularization on species-species associations
We can do the same for the species associations:

```{r,eval=FALSE}
model = sjSDM(Y = com$response, 
              env = linear(data = com$env_weights, formula = ~0+ I(X1^2),lambda = 0.5),
              biotic = bioticStruct(lambda =0.1),
              iter = 50L)
summary(model)
```

Regularization on the inverse covariance matrix (precision) is also supported:
```{r,eval=FALSE}
model = sjSDM(Y = com$response, 
              env = linear(data = com$env_weights, formula = ~0+ I(X1^2),lambda = 0.5),
              biotic = bioticStruct(lambda =0.1, on_diag = FALSE,inverse = TRUE),
              iter = 50L)
summary(model)
```



# Fitting a non-linear (deep neural network) model 
```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
X = com$env_weights
Y = com$response

# three fully connected layers with relu as activation function
model = sjSDM(Y = Y, 
              env = DNN(data = X, formula = ~., hidden = c(10L, 10L, 10L), activation = "relu"), 
              iter = 50L, se = TRUE)
summary(model)
```



The methods for sjSDM() work also for the non-linear model:
```{r,eval=FALSE}
getCov(model) # species association matrix
pred = predict(model) # predict on fitted data
pred = predict(model, newdata = X) # predict on new data
```

Extract and set weights of model:
```{r,eval=FALSE}
weights = getWeights(model) # get layer weights and sigma
setWeights(model, weights)
```

Plot the training history:
```{r,eval=FALSE}
plot(model)
```



# Adding a spatial term
```{r,eval=FALSE}
com = simulate_SDM(env = 3L, species = 5L, sites = 100L)
X = com$env_weights
Y = com$response

XYcoords = matrix(rnorm(200), 100, 2)

model = sjSDM(Y, env = linear(X, ~X1+X2), spatial = linear(XYcoords, ~0+X1:X2))
summary(model)
```


## Regularization on the spatial model:

```{r,eval=FALSE}

model = sjSDM(Y, env = linear(X, ~X1+X2), spatial = linear(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```


## We can also use a non-linear model as a spatial model:

```{r,eval=FALSE}

model = sjSDM(Y, env = linear(X, ~X1+X2), spatial = DNN(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```

## Or for both:
```{r,eval=FALSE}

model = sjSDM(Y, env = DNN(X, ~X1+X2), spatial = DNN(XYcoords, ~0+X1:X2, lambda = 0.4))
summary(model)
```
